Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-21 paper/ms.tex
HODLR comparison in paper

Another GP method that has been used extensively in astronomy is the
hierarchical off-diagonal low rank \citep[HODLR,][]{Ambikasaran:2016} solver.
This method exploits the fact that many commonly used kernel functions produce
``smooth'' matrices to approximately compute the GP likelihood with the
scaling $\mathcal{O}(N\,\log^2 N)$.
This method has the advantage that, unlike \celerite, it can be used with any
kernel function but, in practice, the cost can still prove to be prohibitively
high for multi-dimensional inputs.

Many other approximate methods for scaling GP inference exist and we will make
no attempt to discuss the comparisons exhaustively.

Approximate:
- inducing point methods \citep{Quinonero-Candela:2007}
- KISS-GP \citep{Wilson:2015, Wilson:2015a}
- HODLR \citep{Ambikasaran:2016}

>>>>

Another GP method that has been used extensively in astronomy is the
hierarchical off-diagonal low rank \citep[HODLR,][]{Ambikasaran:2016} solver.
This method exploits the fact that many commonly used kernel functions produce
``smooth'' matrices to approximately compute the GP likelihood with the
scaling $\mathcal{O}(N\,\log^2 N)$.
This method has the advantage that, unlike \celerite, it can be used with any
kernel function but, in practice, the cost can still prove to be prohibitively
high for multi-dimensional inputs.
The proportionality constant in the $N\,\log^2N$ scaling of the HODLR method
is a function of the specific kernel and we find~--~using the \project{george}
software package \citep{Foreman-Mackey:2014, Ambikasaran:2016}~--~that this
scales approximately linearly with $J$ but with substantial overhead for small
models when compared to \celerite.
This means that the HODLR solver can \emph{approximately} evaluate \celerite\
models more efficiently than \celerite\ for large models $J \gg 10$ with a
small dataset where $N$ is smaller than a few thousand.

Many other approximate methods for scaling GP inference exist \citep[see, for
example,][and references therein]{Wilson:2015a} and we will make no attempt to
discuss the comparisons exhaustively.
